Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-09 paper/paper.tex

@@

\paragraph{Benchmark.} We consider 5725 instances, which are available in a simple text format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12 classes.

@@

\section{Per-Instance Algorithm Selection for subgraph isomorphism}

@@

@@

\paragraph{Features.}

@@

\paragraph{Algorithm selection.}

@@

>>>>

The Glasgow algorithm has a parameter, which controls the lengths of paths used
when reasoning about non-adjacent vertices.  In experiments reported in
\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable
compromise -- longer paths lead to prohibitively expensive preprocessing on
larger, denser instances. This is often not the best choice on an instance by
instance basis: sometimes path-based reasoning gives no benefit at all,
sometimes considering only paths of length 2 suffices, occasionally paths of
length 4 are helpful, and even looking at paths of length 3 is relatively
expensive on some graphs.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/ecdf-glasgow}
\caption{Empirical cumulative distribution function for the Glasgow algorithm
with path lengths 1, 2, 3, and 4.}
\label{fig:ecdf-glasgow}
\end{figure}

Figure~\ref{fig:ecdf-glasgow} shows the cumulative distribution function for the
Glasgow algorithm with four different path lengths. As the instances become
harder to solve, longer paths become better. This is what we expect, as more
reasoning is expensive, but potentially increases the reduction of the search
space. In particular, the relative performance of the algorithms is monotone in
the time it takes to solve the instance -- path length 1 is best at first, but
gets worse as the time to solve the instance increases until path length 2 is
better. After that point, length 1 is never better than length 2 again. The same
is true for the other path lengths.

The figure illustrates the potential for portfolios and algorithm selection we
have. There is clearly no path length that dominates throughout. Therefore, we
consider 4 different instances of the Glasgow algorithm here, with path lengths
varying from 1 to 4.

@@

\paragraph{Benchmark.}

We consider 5725 instances, which are available in a simple text
format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These
instances are grouped into 12 classes.

@@

\subsection{Per-Instance Algorithm Selection}

@@

\section{Algorithm Selection Approach}

@@

\subsection{Features}

@@

In addition to computing feature values, the LAD preprocessing is able to detect
whether an instance is unsatisfiable. In this case, it is not necessary to run
an algorithm on the instance.

We apply two static steps to each instance before solving it. First, the
computation of the LAD features can detect unsatisfiability of an instance. If
an instance is identified as unsatisfiable, we do not process it further.
Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
without the overhead of running algorithm selection and avoids potentially
making incorrect solver choices.

Computing the LAD features detects unsatisfiability on 1919 instances. Running
VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
instances that we consider for algorithm selection.

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
\caption{Empirical cumulative distribution function of solvers, virtual best
solvers, and the LLAMA portfolio.}
\label{fig:portfolio-ecdf}
\end{figure}

\begin{table}[t]
\begin{center}
\begin{tabular}{|c||r||r|r|r||r|r|r|r||r|r|}
\hline
$t$ & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{Glasgow}& VBS & $\delta$\\
&&pre&def&new&1&2&3&4&&\\\hline
$10^0$ & 300 &  \cellcolor{blue!25}1126 & 661 & 599 & 386 & 95 & 39 & 12 & 1132 & 6\\\hline
$10^1$ & 1699 & 1858 & 2101 &  \cellcolor{blue!25}2287 & 2061 & 1215 & 455 & 208 & 3388 & 101\\\hline
$10^2$ & 2834 & 1912 & 3376 & 3708 &  \cellcolor{blue!25}3943 & 3344 & 2295 & 1329 & 4636 & 693\\\hline
$10^3$ & 3458 & 1918 & 4233 & 4531 &  \cellcolor{blue!25}4908 & 4763 & 4262 & 3426 & 5170 & 262\\\hline
$10^4$ & 3701 & 1919 & 4888 & 5025 & 5156 &  \cellcolor{blue!25}5253 & 5022 & 4408 & 5332 & 79\\\hline
$10^5$ & 3850 & 1919 & 5108 & 5193 & 5301 &  \cellcolor{blue!25}5377 & 5292 & 5082 & 5434 & 57\\\hline
$10^6$ & 3977 & 1919 & 5244 & 5307 & 5386 &  \cellcolor{blue!25}5452 & 5451 & 5235 & 5500 & 48\\\hline
$10^7$ & 4082 & 1919 & 5336 & 5414 & 5458 & 5517 &  \cellcolor{blue!25}5518 & 5425 & 5567 & 49\\\hline
$10^8$ & 4191 & 1919 & 5423 & 5479 & 5508 &  \cellcolor{blue!25}5561 & 5560 & 5554 & 5608 & 47\\\hline
\end{tabular}
\end{center}
\caption{Number of solved instances at different CPU time limits: Each line displays a time limit
$t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD
preprocessing, default LAD, and new LAD, Glasgow with  the lengths of paths limited to 1, 2, 3 and
4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and
single best if displayed in the last column ($\delta$).\label{expTime}}
\end{table}

Table \ref{expTime} displays the number of instances solved at different CPU time limits, ranging
from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time limit
considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very
quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD
preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2
performs a basic backtracking search and is very fast on easy instances, because it does not compute
expensive invariants. Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2
and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow
1, and 100 (resp. 1000 and 10000) ms for Glasgow 2 (resp. 3 and 4). The Virtual Best Solver (VBS),
which considers the best algorithm for each instance separately, obtains much better results,
showing us that the algorithms have complementary performance. The difference between VBS and single
best (column $\delta$ of Table \ref{expTime}) is more particularly important for rather small CPU
time limits. In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the
limit is 100 ms. In many applications, it is important to have the fastest possible algorithm. For
example, in pattern recognition applications, we often have to solve subgraph isomorphism problems
for a very large number of graphs (in order to find a pattern image in a large database of target
images, for example), so that having an algorithm that is able to solve an instance in 100 ms
instead of 1000 ms makes a big difference. Therefore, it is important to select the best algorithm
for each instance.

Table 3 shows the performance of our algorithm selection approach, compared to the VBS (which is the
upper bound of what an algorithm selection approach can achieve), and the single best solver, at
different CPU time limits (the single best corresponds to results highlighted in blue in Table 1).
Comment Table 3: Mean runtime but also number of solved instances at different CPU time limits.
Logically, LLAMA may solve less instances than single best for small time limits because it spends
time to compute features and choose a solver, but it should be better for larger time limits...

\begin{table}
\begin{tabular}{|l|r|rrrrrrrrr|}
&mean & \multicolumn{9}{c|}{Number of solved instances}\\
&runtime & $10^0$ &  $10^1$ &  $10^2$ &  $10^3$ &  $10^4$ &  $10^5$ &  $10^6$ &  $10^7$ &  $10^8$\\\hline
VBS & 2375.9 & 1132 & 3388 & 4636 & 5170 & 5332 & 5434 & 5500 & 5567 & 5608\\\hline
Single best & 3177.6 & 1126 & 2287 & 3943 & 4908 & 5253 & 5377 & 5452 & 5518 & 5561\\\hline
LLAMA & 2706.3\\\hline
\end{tabular}
\caption{}
\end{table}

Add Figure 1 which compares single best at $10^8$ ms (Glasgow 2) with LLAMA ?
